---
title: "Some difficulties in the `Oarray` package"
author: "Robin K. S. Hankin"
output: html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Oarray_difficulties}
  %\usepackage[utf8]{inputenc}
---


```{r setup, include=FALSE}
options(rmarkdown.html_vignette.check_title = FALSE) 
knitr::opts_chunk$set(echo = TRUE)
library("Oarray")
set.seed(0)
```

This short document gives near-minimal examples showing the problems
of evaluation within accessor methods in the `Oarray` package.  Here
we see an extractor method for the "foo" class, which uses
`match.call()` in the same way as `[.Oarray()` of the `Oarray` package.


```{r defineextractfoo}
`[.foo` <- function(x, ...){
  mc <- match.call()
  fi <- sys.call()

  cat("mc gives: ")
  print(mc)   # NB: cannot cat(mc) as cat() cannot handle language types.

  for(i in 1:6){
    cat(paste("mc[[",i, "]] gives: |",sep=""))
    cat(mc[[i]])
    cat("|\n")
  }
  for(i in 1:6){
    cat(paste("fi[[",i, "]] gives: |",sep=""))
    cat(fi[[i]])
    cat("|\n")
  }

  cat("In the above, the output of print() is given between the two vertical lines.\n")

  for(i in 0:9){
    cat(noquote(paste("eval(mc[[5]],",i,") evaluates to " , eval(mc[[5]],i),"\n",sep="")))
  }

  cat("----\n")
  
  for(i in 0:9){
    cat(noquote(paste("eval(fi[[5]],",i,") evaluates to " , eval(fi[[5]],i),"\n",sep="")))
  }

  print(noquote(paste("mc[[5]]  evaluates to " , eval(mc[[5]],i),sep="")))

  cat("\n\n\n\n\n")
  return(0)
}
```


Above, in function `[.foo()`, `mc` is `match.call()` and we are
interested in `mc[[5]]` because this is `p` in the calls below (such
as `"a[,,p=2,]"`).  

```{r executeextractfoo}
a <- 0
class(a) <- "foo"

p <- 1
a[,,p=2,]   
```

Above, the first element of `mc` is the function call, the second is
`a`, the third and fourth are `<missing>`, the fifth is just 2.  Now
we will try using the value of variable `p`:


```{r executeextractbutusingp}
p <- 1
a[,,p,]
```

Above, we use `p` as a variable (that happens to have the value 1).
So both these examples work as intended.  But, I will give some
examples that do not work as intended, specifically some R idiom in
which symbol `p` exists in multiple environments.  We will create a
function `f()` and access `[.foo()` from within `f()`.  First, we will
pass the value of `p` to `f()`:

```{r fp2}
p <- 1
f <- function(a,p){
    print(p)
    a[,,p,]
}
f(a,p)
```

Above we see that `f()` works as intended: it is picking up the value
of `p` from the global environment (viz, 1).  Now we will try passing
a named argument to `f()`:

```{r fpmultipleenv}
p <- 1
f(a,p=3)
```

Above we want `p` to have the value 3, but all the evaluations of
`mc[[5]]` (which is he symbol `p`) evaluate to 1, its value in the
global environment.  This is not what we want: it is surprising and
misleading.  Now an unnamed argument:


```{r fpmult}
p <- 1
f(a,3)
```

Above we see another surprising result: even though `f()`'s scope has
a value for `p` of 3, `mc[[5]]` nevertheless returns the global value
of 1.  Final example (which gives an error) shows that p might not
exist in all environments that might be accessed:

```{r removep, error=TRUE}
rm(p)
h(p=7)
```
